---
title: "scml_iris"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scml_iris}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE}
library(scml)
library(torch)
library(luz)
```

In this notebook, we show how to train a multiclass classifier and an ordinal regression model on `iris` dataset.

# Train a classifier on iris dataset

A multiclass model is trained below on 200 epoch over `iris` dataset keeping 10% of the data as validation set. The inut dimensionality being low, we limit the input drop out to 10%.


```{r message=FALSE}
fm <- train_delayed_classifier(
  data.matrix(iris[1:4]),
  iris$Species,
  pre_pruning_epoch = 100L,post_pruning_epoch = 100L,
  batch_size = 150L,
  valid_data = 0.1,
  input_dropout_rate = 0.1,
  accelerator = accelerator(cpu=TRUE)
)
```

learning curves
```{r fig.height=6, fig.width=6}
plot(fm)
```

performance on training(+valid) set
```{r eval=FALSE, warning=FALSE}
pred <- predict_delayed(data.matrix(iris[1:4]),fm,accelerator = accelerator(cpu=TRUE))
table(iris$Species,max.col(pred),useNA="i")
```



# Train an ordinal regressor on iris dataset

To train an ordinal regression model we assume the 3 iris labels are ordered (for example it can represent the plant maturity, Baby, Young, Adult).

```{r,message=FALSE,fig.height=3, fig.width=4}
fm <- train_delayed_classifier(
  data.matrix(iris[1:4]),
  iris$Species,
  y_type = "ordinal",
  batch_size = 150L,
  valid_data = 0.1,
  input_dropout_rate = 0.25,
  pre_pruning_epoch = 100L,post_pruning_epoch = 100L,
  accelerator = accelerator(cpu=TRUE)
)
```

learning curves
```{r,message=FALSE,fig.height=3, fig.width=4}
plot(fm)
```

performance on training(+valid) set
```{r,fig.height=4, fig.width=4}
pred <- predict_delayed(data.matrix(iris[1:4]),fm,accelerator = accelerator(cpu=TRUE))
plot(pred~iris$Species)
```







